---
title: "1.4 — The Ricardian View of Trade: Formal One-Factor Model — Practice Problems (Solutions)"
author: "ECON 324 — International Trade"
date: "Tuesday August 25, 2020"
output:
  pdf_document:
    latex_engine: xelatex
    dev: cairo_pdf
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = F)
library(tidyverse)
library(pander)
library(patchwork)
```

Suppose the following facts to set up:

- Home has 100 Laborers
  - Requires 5 workers to make 1wheat
  - Requires 10 workers to make 1 car
  
- Foreign has 200 Laborers
  - Requires 2 workers to make 1 wheat
  - Requires 8 workers to make 1 car

Plot wheat (w) on the horizontal axis and cars (c) on the vertical axis.

**1. Write the equation of each country's production possibilities frontier (PPF).**

Home:
$$\begin{aligned}
l_ww+l_cc&=\bar{L}\\
5w+10c &= 100\\
10c&=100-5w\\
c&=10-0.5w\\
\end{aligned}$$

Foreign:
$$\begin{aligned}
l_w'w+l_c'c&=\bar{L}'\\
2w+8c &= 200\\
8c&=200-2w\\
c&=20-0.25w\\
\end{aligned}$$

\clearpage

**2. Graph each country's PPF.**

```{r}
home_ppf=function(x){10-0.5*x}

home<-ggplot(data = tibble(x=1:10), aes(x=x))+
  stat_function(fun = home_ppf, geom = "line", color = "blue", size = 1)+
  labs(title = "Home",
       x = "Wheat",
       y = "Cars")+
  scale_x_continuous(breaks = seq(0,100,10),
                     limits = c(0,100),
                     expand = c(0,0))+
  scale_y_continuous(breaks = seq(0,100,10),
                     limits = c(0,100),
                     expand = c(0,0))+
  theme_classic(base_family = "Fira Sans Condensed", base_size=16)

foreign_ppf=function(x){20-0.25*x}

foreign<-ggplot(data = tibble(x=1:10), aes(x=x))+
  stat_function(fun = foreign_ppf, geom = "line", color = "red", size = 1)+
  labs(title = "Foreign",
       x = "Wheat",
       y = "Cars")+
  scale_x_continuous(breaks = seq(0,100,10),
                     limits = c(0,100),
                     expand = c(0,0))+
  scale_y_continuous(breaks = seq(0,100,10),
                     limits = c(0,100),
                     expand = c(0,0))+
  theme_classic(base_family = "Fira Sans Condensed", base_size=16)

home+foreign
```

\vspace{1in}

**3. Which country has an *absolute* advantange in producing each good? Why?**

Foreign has an absolute advantage in producing wheat, since it requires only 2 workers to produce 1 wheat, compared to Home's 5 workers. Also see that Foreign can produce more wheat (80 units) than Home (20 units) if it dedicated all of its resources to wheat production — the horizontal intercept.

Foreign also has an absolute advantage in producing cars, since it requires only 8 workers to produce 1 car, compared to Home's 10 workers. Also see that Foreign can produce more cars (20 units) than Home (10 units) if it dedicated all of its resources to car production — the vertical intercept.

\clearpage

**4. Find each country's opportunity cost of producing each good.**

| Country | 1 Wheat | 1 Car |
|:-------|--------:|------:|
| Home | 0.5 cars | **2 wheat** |
| Foreign | **0.25 cars** | 4 wheat |

Each country's opportunity cost of wheat is the slope of its PPF: the relative price of wheat, or, the number of cars needed to be given up to produce 1 additional unit of wheat.

Each country's opportunity cost of cars is the *inverse* of the slope of its PPF: the relative price of cars, or, the number of wheat needed to be given up to produce 1 additional car.

\vspace{1in}

**5. Which country has a *comparative* advantage in producing each good? Why?**

Home has a comparative advantage in producing cars since it has a lower opportunity cost in car production. It only gives up 2 wheat vs. Foreign gives up 4 wheat per car. 

Foreign has a comparative advantage in producing wheat since it has a lower opportunity cost in wheat production. It only gives up 0.25 cars vs. Home gives up 0.5 cars per 1 wheat. 

\vspace{1in}

**6. Which country should specialize in producing which good?**

Each country should specialize in producing the good for which they have the lower opportunity cost: Home should specialize in cars and Foreign should specialize in wheat.

\vspace{1in}

**7. What will the range of possible terms of trade be?**

The price of each good (in terms of the other) must fall between the opportunity cost of each country to produce that good. Thus is because a country will only buy (import) a good if it can get a price *cheaper* than it costs it to produce at home, and a country will only sell (export) a good if it can get a price *higher* than its cost of production:

Simply look down each column for each good in the table above.

\begin{center}
0.25 cars $< p_W <$ 0.5 cars \\
2 wheat $< p_C <$ 4 wheat
\end{center}